---
title: "MIT VAR Example"
output: html_notebook
---

From https://ocw.mit.edu/courses/mathematics/18-s096-topics-in-mathematics-with-applications-in-finance-fall-2013/case-studies/MIT18_S096F13_CaseStudy5.pdf

### Load data
```{r}
getSymbols(c("UNRATE", "FEDFUNDS", "TB3MS", "CPIAUCSL", "M1SL", "GDPDEF", "GDP", "GPDI", "TWEXBMTH"), src="FRED")
SP500 <- getSymbols("^GSPC", src="yahoo", auto.assign=F)
```

### 1.3 Ordinary and Partial Autocorrelations of Reduced Set
```{r}
# Consider focusing on 3 variables
ymat0<-merge(UNRATE, FEDFUNDS, CPIAUCSL)
ind.quarterly0<-1*(is.na(ymat0[,3])==FALSE)
sum(ind.quarterly0)
dim(ymat0)
```

```{r}
ymat00<-ymat0[which(ind.quarterly0==1),]
head(ymat00)
```

```{r}
par(mfcol=c(3,1))
plot(ymat00[,1],main=dimnames(ymat00)[[2]][1])
plot(ymat00[,2],main=dimnames(ymat00)[[2]][2])
plot(ymat00[,3],main=dimnames(ymat00)[[2]][3])
```

###
```{r}
 ymat00.0<-window(ymat00
                  , start = as.Date("1960-01-01")
                  , end = as.Date("2000-12-31"))
dim(ymat00.0)
head(ymat00.0)
```

```{r}
acf(ymat00.0, lag.max=10)
```

```{r}
acf(ymat00.0, type="partial", lag.max=10)
```

### 1.4 Vector Autoregressive (VAR) Model of Reduced Set
```{r}
 # The function VARselect() is from the package vars; see Pfaff(2008).
# This function identifies the optimal VAR(p) order p.
ymat00.0.VAR.const<-VARselect(ymat00.0, lag.max=12, type="const")
# Print out the VAR order identified by different information criteria
ymat00.0.VAR.const$selection
```

```{r}
# Fit the VAR model corresponding to the Schwarz Criterion (SC) which is the BIC
ymat00.0.VAR.const.0<-VAR(ymat00.0, p=ymat00.0.VAR.const$selection[3],type="const")
options(show.signif.stars=FALSE)
summary(ymat00.0.VAR.const.0)
```

### 1.5 Impulse Response Functions for a Fitted VAR(p) Model
The impulse response function measure the impact of a unit innovation (impulse)
in a given variable on all the dependent variables in the VAR model.
```{r}
plot(irf(ymat00.0.VAR.const.0, impulse="UNRATE"))

# When unemployment rises:
# the Federal Funds rate is projected to decline
# (consistent with Federal Reserve Policy)
#
# the CPI decreases (lower employment results in less
# pressure to increase consumer prices)
```
```{r}
plot(irf(ymat00.0.VAR.const.0, impulse="FEDFUNDS"))

# When the Fed Funds rate increases:
#
# The Unemployment rate tends to increase;
# so reducing the Fed Funds rate would tend to reduce unemployment
#
# The CPI increases; increases in the Fed Funds rate are
# associated with increase in CPI over future quarters
```
```{r}
plot(irf(ymat00.0.VAR.const.0, impulse="CPIAUCSL"))
```

When the CPI increases

The Federal Funds rate tends to increase over subsequent quarters.
This is consistent with Federal Reserve policy of raising
interest rates to control for inflation.

### 1.6 Ordinary and Partial Autocorrelations of Differenced Series
```{r}
ymat000.0<-na.omit(diff(ymat00.0))
acf(ymat000.0, lag.max=10)
```

```{r}
acf(ymat000.0, type="partial", lag.max=10)
```

### 1.7 Vector Autoregressive (VAR) Model with Differenced Series
```{r}
# The function VARselect() is from the package vars; see Pfaff(2008).
# This function identifies the optimal VAR(p) order p.
ymat000.0.VAR.const<-VARselect(ymat000.0, lag.max=12, type="const")
# Print out the VAR order identified by different information criteria
ymat000.0.VAR.const$selection
# Fit the VAR model corresponding to the Schwarz Criterion (SC) which is the BIC
ymat000.0.VAR.const.0<-VAR(ymat000.0, p=ymat000.0.VAR.const$selection[3],type="const")
options(show.signif.stars=FALSE)
summary(ymat000.0.VAR.const.0)
```

### 1.8 Impulse Response Functions for VAR(p) Fit of Differenced Series
```{r}
plot(irf(ymat000.0.VAR.const.0, impulse="UNRATE"))
```

```{r}
plot(irf(ymat000.0.VAR.const.0, impulse="FEDFUNDS"))
```

```{r}
plot(irf(ymat000.0.VAR.const.0, impulse="CPIAUCSL"))
```

Interpreting the impulse response functions for the VAR model of the differenced series, we note:
• When unemployment increases, the Fed Funds rate tends to decrease over
subsequent quarters, consistent with Federal Reserve policies (i.e., stimulating economic growth and employment with lower interest rates).
• When the Fed Funds rate increases, there is a modest increase in inflation
(CPIA). This is consistent with the Fed raising rates to control inflation
which tends to persist for several quarters (note the high 3-rd quarter lag
partial autocorrelation in CPIAUCSL).
• When inflation (CPIAUCSL) increases, unemployment tends to rise modestly, and the Fed Funds rate tends to increase.

References
Bernard Pfaff (2008). VAR, SVAR and SVEC Models: Implementation With
R Package vars, Journal of Statistical Software 27(4). URL http://www.jstatsoft.org/v27/i04/.
Robert Litterman (1979). Techniques of Forecasting Using Vector Autoregressions. Working Paper # 115, Federal Reserve Bank of Minneapolis.
Christopher Sims (1989). A Nine Variable Probabilistic Macroeconomic Forecasting Model. Discussion Paper 14, Federal Reserve Bank of Minneapolis.
