---
title: "R Notebook"
output: html_notebook
---

# Clear rates data
```{r}
caim::clr()
library(caim)
library(data.table)
```

# Set up Parameters
```{r}
CURVE_COUNTRIES <- c("US", "CA", "DE", "FR", "IT", "UK", "JP", "CN")
CURVE_SECTORS <- c("GOVT")
CURVE_MATURITIES <- c(0.25, 1, 2, 3, 5, 7, 10, 30)
KEY_MATURITIES <- c(0.25, 1, 2, 5, 10, 30)

NUM_SIMS <- 5000
NUM_PERIODS_IN_SIM <- 12 # 12 = annual, etc.
SIM_DECAY_FACTOR <- 1
SIM_BLOCK_SIZE <- 3

RATES_INFO_FILE <- "../../bb_data/rates_info.csv"
TOT_RET_INFO_FILE <- "../../bb_data/tot_ret_info.csv"
DATA_FILE <- "../../bb_data/bb_data.csv"

YIELD_PATH_CHART_COUNTRY <- "US"
YIELD_PATH_CHART_CLASS <- "GOVT"
YIELD_PATH_CHART_MATURITY <- 2
YIELD_PATH_CHART_NUM_SERIES <- 250
```

# Load rates data from data directory
```{r}
rates_info <- fread(RATES_INFO_FILE)[
  country %in% CURVE_COUNTRIES & class %in% CURVE_SECTORS
  , .(country, class, maturity, bb_ticker)
]

# extract daily rates data
rates_data_d <- rates_info[
  fread(DATA_FILE)
  ,
  # join on bb_ticker field
  , on=.(bb_ticker)
  # inner join
  , nomatch=NULL
][
  # filter on field
  field == "PX_LAST"
][
  , .(country
      , class
      , maturity
      , date = lubridate::as_date(date)
      # , year = lubridate::year(lubridate::as_date(date))
      # , month = lubridate::month(lubridate::as_date(date))
      , yield = value
      )
][
  # set key and sort
  , .SD, key = .(country, class, maturity, date)
]

# extract monthly rates data
rates_data_m <- caim::monthly_data(rates_data_d, c("country", "class", "maturity"))[
  , .SD
  , keyby = .(country, class, maturity, date)
]

# find valid dates
rates_date_range <- caim::date_range(
  rates_data_m[maturity %in% KEY_MATURITIES],
  c("country", "class", "maturity")
)
# find first date where all curves and key maturities have data
first_date <- max(rates_date_range$min_date)

# KLUDGE: the above needs reworking, ta accept 1 yr short end for China 
first_date <- lubridate::as_date("2005-06-30")

# find last date where all curves and key maturities have data
last_date <- min(rates_date_range$max_date)
# extract relevant date range
rates_data_m <- rates_data_m[
  date >= first_date & date <= last_date
]

rates_date_range[]
rates_data_m[]
```
```{r}
# fix tot_ret_data in bb_data
# bb_data <- fread("../../bb_data/bb_data.csv")
# tot_ret_tickers <- tot_ret_info$bb_ticker
# bb_data[bb_ticker %in% tot_ret_tickers, date := as.character(lubridate::mdy(date))]
# fwrite(bb_data, "../../bb_data/bb_data.csv")
```

#### Load total return asset data
```{r}
# read tot_ret_info file
tot_ret_info <- fread(TOT_RET_INFO_FILE)

# extract daily total return data
tot_ret_data_d <- tot_ret_info[
  fread(DATA_FILE)
  ,
  # join on bb_ticker field
  , on=.(bb_ticker)
  # inner join
  , nomatch=NULL
][
  # filter on field
  field == "PX_LAST"
][
  , .(
    class
    , name
    , date
    , value
    , ltcma
    , class_id
    , asset_id
  )
]
setkeyv(tot_ret_data_d, c("class_id", "asset_id", "date"))

# extract monthly total return data
tot_ret_data_m <- caim::monthly_data(tot_ret_data_d, c("class", "name"))
setkeyv(tot_ret_data_m, c("class_id", "asset_id", "date"))
# calculate historic monthly returns
tot_ret_data_m[ , hist_return := value / shift(value), by=.(class_id, asset_id)]
# detrend historic returns and add forward-looking return expectation
tot_ret_data_m[!is.na(hist_return) , return := hist_return - as.numeric(caim::mean_exp(hist_return - 1, SIM_DECAY_FACTOR)), by=.(class_id, asset_id)]
# + (ltcma / 12)

tot_ret_stats_m <- tot_ret_data_m[
  !is.na(return)
  , .(
    mean = mean(return - 1) * 12
    , vol = sd(return - 1) * sqrt(12)
    , freqless0 = caim::freq_below(return - 1, 0)
    , var95 = caim::var_emp(return - 1, 0.95)
    , cvar95 = caim::cvar_emp(return - 1, 0.95)
  )
  , by=.(class_id, asset_id)
]

tot_ret_data_m[]
tot_ret_stats_m[]
```


#### Linear interpolated yields
```{r}
# find every combination of country, class, maturity and dte
points_all <- data.table(expand.grid(
  stringsAsFactors = F
  , country = sort(unique(rates_data_m$country))
  , class = sort(unique(rates_data_m$class))
  , maturity = sort(unique(rates_data_m$maturity))
  , date = lubridate::as_date(sort(unique(rates_data_m$date)))
))
setkeyv(points_all, c("country", "class", "maturity", "date"))
points_all[]

# find min and max maturities with valid data for each date for each curve
maturity_range <- rates_data_m[ 
  , .(
    mat_min = min(maturity)
    , mat_max = max(maturity)
    )
  , keyby=.(country, class, date)
]

# joins rates_data_m with points_all, will have NAs where there is no data;
#   these will be filled in with interpolated data
rates_data <- rates_data_m[points_all][maturity_range, on=.(country, class, date)]
# linear interpolation. rule=2 will extend min and max values beyond valid maturity ranges
rates_data[ , yield_approx := approx(x = maturity, y = yield, xout = maturity, rule = 2)$y, by=.(country, class, date)]
# spline interpolation. useful because method = "natural" will linear extrapolate values outside valid maturity range
rates_data[ , yield_spline := spline(x = maturity, y = yield, xout = maturity, method="natural")$y, by=.(country, class, date)]
# use linear interpolated data - will keep these values for all observations within valid maturity range
rates_data[ , yield_lin := yield_approx]
# for extrapolated data, use average of constant min/max and linear extrapolation. this should moderate crazy extrapolated values
rates_data[maturity < mat_min | maturity > mat_max, yield_lin := mean(c(yield_approx, yield_spline)), by=.(country, class, maturity, date)]

rates_data[]
```

# filter to use preferred curve points
```{r}
curve_rates_data <- rates_data[maturity %in% CURVE_MATURITIES]
# once we're happy with the calculation components of curve_rates_data, clean up columns
curve_rates_data <- curve_rates_data[ , .(yield=yield_lin), key=.(country, class, date, maturity)]

curve_rates_data[]
```
# calculate historic performance at curve points
```{r}
hist_rates_obs <- copy(curve_rates_data)
# calculate duration for each curve point at each point in time. duration is time-varying, not constant, depending on yields
hist_rates_obs[ , mod_dur := caim::modified_duration(yield / 100, yield / 100, maturity, ifelse(country=="US", 2, 1))]
# calculate "buy yield" - previous month's yield at this maturity point
hist_rates_obs[ , yield_buy := shift(yield), by=.(country, class, maturity)]
# find interpolated "sell yields", with maturity = maturity - 1 month (we've held the security for a month)
hist_rates_obs[ , yield_sell_approx := approx(x = maturity, y = yield, xout = maturity - 1/12, rule = 2)$y, by=.(country, class, date)]
# use spline interpolation to get linear extrapolation outside maturity range
hist_rates_obs[ , yield_sell_spline := spline(x = maturity, y = yield, xout = maturity - 1/12, method = "natural")$y, by=.(country, class, date)]
# use linear interpolation values (only those within maturity range will be kept)
hist_rates_obs[ , yield_sell := yield_sell_approx]
# use average of min_yield and spline extrapolation for maturities outside valid maturity range
mat_min <- min(hist_rates_obs$maturity)
hist_rates_obs[(maturity - 1/12) < mat_min , yield_sell := mean(c(yield_sell_approx, yield_sell_spline)), by=.(country, class, maturity, date)]
# if happy with yield_sell_approx, yield_sell_spline and yield_sell, clean up
hist_rates_obs <- hist_rates_obs[ , !c("yield_sell_approx", "yield_sell_spline")]
# coupon income = previous month's yield / 12
hist_rates_obs[ , coup_inc := yield_buy / 1200]
# price = par bond bought at previous month's yield, sold at this month's maturity - 1 month yield
hist_rates_obs[ , price_inc := caim::bond_price(yield_sell / 100, yield_buy / 100, maturity - 1/12, ifelse(country=="US", 2, 1))/100]
# total return: price income + coupon income
hist_rates_obs[ , tot_ret := price_inc + coup_inc]

hist_rates_obs[]
```
# simulations
```{r}
# add date index so sims can access it
sim_rates_data <- copy(curve_rates_data)[ 
  , date_ix := 0:(.N - 1)
  , by=.(country, class, maturity)
][
  , .(yield)
  , key=.(country, class, date_ix, date, maturity)
]
# previous yield for troubleshooting
sim_rates_data[ , yield_prev := shift(yield), by=.(country, class, maturity)]
# yield changes in bps
sim_rates_data[ , yield_chg := yield - shift(yield), by=.(country, class, maturity)]
# detrend historic yield changes
sim_rates_data[!is.na(yield_chg) , yield_chg := yield_chg - as.numeric(caim::mean_exp(yield_chg, SIM_DECAY_FACTOR)), by=.(country, class, maturity)]
# yield changes as pct of previous value
sim_rates_data[ , yield_chg_pct := yield / shift(yield), by=.(country, class, maturity)]

num_obs <- max(sim_rates_data$date_ix)
sim_ix <- caim::sim_period_ix(NUM_SIMS, NUM_PERIODS_IN_SIM, num_obs, SIM_BLOCK_SIZE, caim::d_exp(SIM_DECAY_FACTOR, num_obs))

print("constructing sim observations")
print(system.time(sim_rates_obs <- rbindlist(list(
  sim_rates_data[
    date_ix==num_obs
    , .(per=0, country, class, date_ix, date, maturity, yield_chg=yield)
  ][
    data.table(sim=1:NUM_SIMS, per=0)
    , on=.(per), allow.cartesian=T
  ]
  , copy(sim_rates_data[
    , .(yield_chg)
    , key=.(country, class, date_ix, date, maturity)
  ])[
    sim_ix, on=.(date_ix=ix)
    , allow.cartesian=T
  ]
), use.names=T)[
  , .(sim, per, country, class, maturity, date_ix, date, yield_chg)
]
))
print("setting key")
print(system.time(
  setkeyv(sim_rates_obs, c("sim", "country", "class", "maturity", "per"))
))
print("calculating yield path")
print(system.time(
  sim_rates_obs[ , yield := cumsum(yield_chg), by=.(country, class, sim, maturity)]
))
print("adjusting yields to minimum bounds")
print(system.time(
  sim_rates_obs[ , yield_adj := max(yield, -0.5), by=.(country, class, sim, per, maturity)]
))
print("calculating buy yields")
print(system.time(
  sim_rates_obs[ , yield_buy := shift(yield_adj), by=.(country, class, sim, maturity)]
))
print("calculating sell yields")
print(system.time(
# find interpolated "sell yields", with maturity = maturity - 1 month (we've held the security for a month)
  sim_rates_obs[ , yield_sell := approx(x = maturity, y = yield_adj, xout = maturity - 1/12, rule = 2)$y, by=.(country, class, sim, per)]
))
print("calculating coupon income")
print(system.time(
  sim_rates_obs[ , coup_inc := yield_buy / 1200]
))
print("calculating price income")
print(system.time(
  sim_rates_obs[ , price_inc := caim::bond_price(yield_sell / 100, yield_buy / 100, maturity - 1/12, ifelse(country=="US", 2, 1))/100]
))
print("calculating total return")
print(system.time({
  sim_rates_obs[ , tot_ret := price_inc + coup_inc]
  sim_rates_obs[is.na(tot_ret), tot_ret := 1.0]
}))

sim_rates_returns_ann <- sim_rates_obs[ , .(return = prod(tot_ret, na.rm=T) - 1), by=.(country, class, maturity, sim)]
sim_rates_stats_ann <- sim_rates_returns_ann[ 
  , .(
    mean = mean(return)
    , vol = sd(return)
    , freqless0 = caim::freq_below(return, 0)
    , var95 = caim::var_emp(return, 0.95)
    , cvar95 = caim::cvar_emp(return, 0.95)
  )
  , by=.(country, class, maturity)
]

sim_rates_data[]
sim_rates_obs[]
sim_rates_stats_ann[]
```
#### Total return asset simulations
```{r}
# add date index so sims can access it
sim_tot_ret_data <- copy(tot_ret_data_m)[ 
  , date_ix := 0:(.N - 1)
  , by=.(class_id, asset_id)
][
  , .(class, name, ltcma, return)
  , key=.(class_id, asset_id, date_ix, date)
]

print("constructing sim observations")
print(system.time(sim_tot_ret_obs <- copy(sim_tot_ret_data[
    , .(class, name, ltcma, return)
    , key=.(class_id, asset_id, date_ix, date)
  ])[
    sim_ix, on=.(date_ix=ix)
    , allow.cartesian=T
  ][
  , .(sim, per, class_id, asset_id, class, name, ltcma, date_ix, date, return)
]
))
print("setting key")
print(system.time(
  setkeyv(sim_tot_ret_obs, c("sim", "class_id", "asset_id", "per"))
))

# calculate expected volatilities (return expectations not added in yet)
sim_tot_ret_returns_ann <- copy(sim_tot_ret_obs[ , .(return = prod(return, na.rm=T) - 1), by=.(class_id, asset_id, sim, class, name, ltcma)])
# for volatile assets, simulations could have wild asymmetric outcomes and means might be significantly different than 0, so detrend
sim_tot_ret_returns_ann <- sim_tot_ret_returns_ann[ , return := ltcma + return - mean(return), by=.(class_id, asset_id, class, name)]

sim_tot_ret_stats_ann <- sim_tot_ret_returns_ann[ 
  , .(
    mean = mean(return)
    , vol = sd(return)
    , freqless0 = caim::freq_below(return, 0)
    , var95 = caim::var_emp(return, 0.95)
    , cvar95 = caim::cvar_emp(return, 0.95)
  )
  , by=.(class_id, asset_id)
]

sim_tot_ret_data[]
sim_tot_ret_obs[]
sim_tot_ret_returns_ann[]
sim_tot_ret_stats_ann[]

```


#### Yield paths and returns
```{r}
# yield paths
yield_paths_x <- 0:NUM_PERIODS_IN_SIM
yield_paths <- dcast(
  sim_rates_obs[
    country == YIELD_PATH_CHART_COUNTRY 
      & class == YIELD_PATH_CHART_CLASS
      & maturity == YIELD_PATH_CHART_MATURITY
      & sim <= YIELD_PATH_CHART_NUM_SERIES
    , .(
      per
      , sim
      , yield_adj
    )
  ]
  , per ~ sim, value.var = "yield_adj"
)
x <- yield_paths$per
y <- yield_paths[ , !c("per")]
plot(
  range(x), range(y)
  , type="n"
  , bty="n"
)
abline(h=0)
for (i in 1:YIELD_PATH_CHART_NUM_SERIES)
  lines(
    x, unlist(y[ , ..i])
    , col=caim::caim_colors(i, 0.15)
    , lwd=2
  )

# return paths
return_paths <- dcast(
  sim_rates_obs[
    country == YIELD_PATH_CHART_COUNTRY 
      & class == YIELD_PATH_CHART_CLASS
      & maturity == YIELD_PATH_CHART_MATURITY
      & sim <= YIELD_PATH_CHART_NUM_SERIES
    , .(
      per
      , ret = cumprod(tot_ret) - 1
    )
    , by = .(country, class, sim, maturity)
  ]
  , per ~ sim, value.var = "ret"
)
x <- return_paths$per
y <- return_paths[ , !c("per")]
plot(
  range(x), range(y)
  , type="n"
  , bty="n"
)
abline(h=0)
for (i in 1:YIELD_PATH_CHART_NUM_SERIES)
  lines(
    x, unlist(y[ , ..i])
    , col=caim::caim_colors(i, 0.15)
    , lwd=2
  )

# sim_rates_obs[
#     country == YIELD_PATH_CHART_COUNTRY 
#       & class == YIELD_PATH_CHART_CLASS
#       & maturity == YIELD_PATH_CHART_MATURITY
#       & sim <= YIELD_PATH_CHART_NUM_SERIES
#       # & per > 0
#     , .(
#       per
#       , ret = cumprod(tot_ret) - 1
#     )
#     , by = .(country, class, sim, maturity)
#   ]
```

# ------------------------------------------------------------------------------
# DEPRECATED
# ------------------------------------------------------------------------------

#### calculate yield changes
```{r}
# # yield changes in bps
# rates_data[ , yield_chg := yield_lin - shift(yield_lin), by=.(country, class, maturity)]
# # yield changes as pct of previous value
# rates_data[ , yield_chg_pct := yield_lin / shift(yield_lin), by=.(country, class, maturity)]
# 
# rates_data[]
```


#### Linear interpolated yields
```{r}
# # ensure every country class and date has every available maturity
# fis <- data.table(expand.grid(stringsAsFactors = F
#   , country = unique(rates_data_m$country)
#   , class = unique(rates_data_m$class)
#   # , date = unique(rates_data_m$date)
#   , maturity = unique(rates_data_m$maturity)
# ))
# setkeyv(fis, c("country", "class", "maturity"))
# fis[]


  
# # some/many observations will be NA
# a <- unique(data.table(rates_data_m[ , .(country, class, date)]))
# b <- unique(data.table(rates_data_m[ , .(class, maturity)]))
# matlist <- a[b, on=.(class), allow.cartesian=T]
# rates_data_m <- rates_data_m[
#   matlist
#   , on=.(country, class, maturity, date)
# ][ 
#   , .SD , key=.(country, class, maturity, date)
# ]

# # fill in missing yields with interpolated data
# # table with !is.na(yield)
# yields_valid <- rates_data_m[!is.na(yield)][, c("mat", "yld") := .(maturity, yield)]

# yields_valid[
#   , c("mat_min", "mat_min_hi", "mat_max_lo", "mat_max") :=
#     .(
#       min(mat)
#     , mat_min_hi = sort(mat)[2]
#     , mat_max_lo = sort(mat)[length(sort(mat)) - 1]
#     , mat_max = max(mat)
#     )
#   , by=.(country, class, date)
# ]

# yields_range <- yields_valid[ 
#   , .(
#     mat_min = min(mat)
#     , mat_min_hi = sort(mat)[2]
#     , mat_max_lo = sort(mat)[length(sort(mat)) - 1]
#     , mat_max = max(mat)
#     )
#   , keyby=.(country, class, date)
# ][
#   yields_valid
#   , yld_min := yld
#   , on=.(country, class, mat_min=mat, date)
# ][
#   yields_valid
#   , yld_min_hi := yld
#   , on=.(country, class, mat_min_hi=mat, date)
# ][
#   yields_valid
#   , yld_max_lo := yld
#   , on=.(country, class, mat_max_lo=mat, date)
# ][
#   yields_valid
#   , yld_max := yld
#   , on=.(country, class, mat_max=mat, date)
# ]

# # table mapping to <=
# yields_lo <- yields_valid[
#   rates_data_m
#   , .(country, class, maturity, date, m_lo=mat, y_lo=yld)
#   , on=.(country, class, date, maturity)
#   , roll=T
# ]
# [
#   is.na(m_lo)
# ][
#   yields_range
#   , c("m_lo", "y_lo") :=
#     .(
#       mat_min
#       , yld_min
#     )
#   , on=.(country, class, date)
# ]

# # table mapping to >=
# yields_hi <- yields_valid[
#   rates_data_m
#   , .(country, class, maturity, date, m_hi=mat, y_hi=yld)
#   , on=.(country, class, date, maturity)
#   , roll=-Inf
# ]
# [
#   is.na(m_hi)
# ][
#   yields_range
#   , c("m_hi", "y_hi") :=
#     .(
#       mat_max
#       , yld_max
#     )
#   , on=.(country, class, date)
# ]
# 
# yields_lin <- yields_hi[
#   yields_lo
#   , .(
#     country
#     , class
#     , maturity
#     , date
#     # , yield_lin = ifelse(m_hi0 == m_lo0
#                          # , y_hi0
#                          # , y_lo0 + (maturity - m_lo0) * (y_hi0 - y_lo0) / (m_hi0 - m_lo0)
#                          # )
#     , m_lo
#     , m_hi
#     , y_lo
#     , y_hi
#   )
#   , on=.(country, class, maturity, date)
# ]

# [
#   is.na(m_lo)
#   , c("m_lo", "y_lo", "m_hi") := 
#     .(m_hi, y_hi, )
# ][
#   is.na(m_hi)
#   , c("m_hi", "y_hi") := 
#     .(m_lo, y_lo)
# ]



# [
#   # lop off "tails" of maturities
#   # kludge until linear interp algo extends beyond min and max
#   maturity >= 0.25 & maturity <= 30
# ][
#   # include only key maturities
#   maturity %in% CURVE_MATURITIES
# ][
#   , .(yield = yield_lin)
#   , key=.(country, class, maturity, date)
# ][
#   , ix := 1:.N
#   , by=.(country, class, maturity)
# ][
#   , .SD
#   , key=.(country, class, ix, maturity, date)
# ]
# 
# yields_lin[]
```

#### Generate curves data
```{r}
# # curve_data <- dcast(rates_data_m, country + class + date ~ maturity, value.var="yield")
# # curve_mats <- sort(unique(rates_data_m$maturity))
# curve_data <- dcast(yields_lin, ix + date ~ country + class + maturity, value.var="yield")
# curve_changes <- curve_data[ , lapply(.SD, function(x) x - shift(x)), .SDcols=!c("ix", "date")]
```

